\documentclass{article}
\usepackage{geometry}
\geometry{a4paper, 
margin=1in}
\usepackage{longtable}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{graphicx} % Required for including images
\usepackage{subcaption}  % Required for subfigures
\usepackage{tcolorbox}  % For colored boxes
% \geometry{margin=1in}
\usepackage{booktabs}    % \toprule, \midrule, \bottomrule
\usepackage{caption}     % if you need \caption* or extra caption control
\usepackage{longtable}   % if your table might span multiple pages
\usepackage{array}       % for more flexible column specs (e.g. >{\ttfamily}l)
\usepackage{multirow}
\usepackage{tabularx}
\usepackage{array}
\usepackage{multirow}
\usepackage{booktabs}

% \title{Zoobenthic Community Indicators of Sediment Contamination}

% \author{Feng Gu}

\begin{document}
% \maketitle

\begin{titlepage}
\centering
\vspace*{2cm}

{\LARGE\bfseries Zoobenthic Community Indicators of Sediment Contamination\par}
\vspace{1.2cm}

{\Large\textbf{Research Proposal}\par}
\vspace{1.5cm}

{\large\textbf{Author:} Feng Gu\par}
\vspace{0.5cm}

{\large\textbf{Program and Institution:}\par}
\vspace*{.2cm}
{\large Master of Science (M.Sc.) in Data Science\\
\vspace*{.2cm}
Thompson Rivers University (TRU), Kamloops, British Columbia, Canada\par}
\vspace{1.5cm}

{\Large\textbf{Supervisory Committee:}\par}
\vspace*{.2cm}
{\large
Dr. Jabed H. Tomal (Supervisor) — Department of Mathematics and Statistics, TRU\\
\vspace*{.2cm}
Dr. Yue Zhang (Co-supervisor) — Department of Mathematics and Statistics, TRU\\
\vspace*{.2cm}
Prof. Jan Ciborowski (Committee Member) — Department of Biological Sciences, University of Calgary
\par}

\vfill
{\large October 2025\par}

\end{titlepage}


\clearpage

\begin{center}
    {\Large \textbf{Abstract}}
\end{center}

This proposal reviews the rationale and analytical framework of Zoobenthic Community Indicators (ZCI)
of Sediment Contamination in aquatic ecosystems. 
Building upon this foundation, the author outlines the research objectives and discusses
the key methodologies to be applied.
Preliminary results are produced based on the framework described in the \textit{Methodology} section,
demonstrating the feasibility of the proposed approach 
and supporting the achievability and significance of the research objectives.  

This proposal comprises eight main sections. 
The \textbf{Introduction} presents the research topic, describes the general relationships among its fundamental components and 
introduces the study system. 
The \textbf{Research Objectives} section defines the specific goals and principal questions to be addressed in the project. 
The \textbf{Data Description} outlines the dataset characteristics, including data types, sources, collection methods, and variable definitions. 
The \textbf{Literature Review} synthesizes relevant research on the close study objectives and their analytical approaches.
The \textbf{Methodology} section presents the overall analytical framework and discusses specific techniques, with several figures provided 
to illustrate the idealized workflow. 
The \textbf{Preliminary Results} section reports initial outcomes from implementing key methodological steps, demonstrating both 
the feasibility and scalability of the proposed framework. 
The \textbf{Practical Implementation Plan} describes the computational strategies supporting the research objectives, including 
the structure of the codebase and choice of programming languages. 
The \textbf{Supervisory Dissolution} section summarizes supervision conditions and acknowledgements, 
and the \textbf{Timeline} outlines a provisional schedule with major milestones for completion of each research phase.  


\clearpage
\tableofcontents

\clearpage
\input{sections/CusCommands.tex} % Custom commands for convenience
\input{sections/Introduction.tex}
\clearpage
\input{sections/ResearchObjectives.tex}
\input{sections/DataDescription.tex}
\input{sections/LiteratureReview.tex}
\input{sections/Methodology.tex}
\input{sections/PreliminaryResults.tex}
\input{sections/subsectionofPreliminaryResults.tex}
\clearpage
\input{sections/ImplementPlan.tex}
\clearpage
\input{sections/SupervisoryDissolution.tex}
\clearpage
\input{sections/Timeline.tex}

\clearpage
\bibliographystyle{plain}  % or use unsrt if you want citations in order of appearance
\bibliography{sections/Reference}   % your .bib file, without .bib extension

\clearpage
\input{sections/Appendix.tex}

\end{document}

