\section{Timeline(temporary)}

The below months are in 2025, each enumerated body in each month is the 
1st, 2nd, 3rd and 4th week of that month.

\begin{itemize}
    \item August
        \begin{enumerate}
            \item Make Preliminary Analysis on the raw data, explore the achievability of the general framework.
            \item Make adjustments on the previous proposal based on the validated preliminary results.
            \item Finish the proposal, check and submit the proposal.
            \item Prepare and study the relevant theoretical foundations, about the math, statistical and coding tools:
            for example: Discriminant Function and Ordination method and Project management.
        \end{enumerate}
    \item September
        \begin{enumerate}
            \item To finish and test the \textbf{\textit{data cleaning, operation and transformation modules}}, 
            design and prepare the \textbf{\textit{PCA sediment contamination module}}.
            \item To finish and test the \textbf{\textit{PCA sediment contamination module}},
            design and prepare the \textbf{\textit{Cluster Analysis module}}.
            \item To finish and test the \textbf{\textit{Cluster Analysis module}},
            design and prepare the \textbf{\textit{Discriminant Function Analysis module}}.
            \item To finish and test the \textbf{\textit{Discriminant Function Analysis module}},
            design and prepare the \textbf{\textit{Ordination-ZCI constructor module}}. 
        \end{enumerate}
    \item October
        \begin{enumerate}
            \item To finish and test the \textbf{\textit{Ordination-ZCI constructor module}},
            design and prepare the \textbf{\textit{Piecewise Quantile Regression module}}.
            \item To finish and test the \textbf{\textit{Piecewise Quantile Regression module}},
            design and prepare the \textbf{\textit{Degradation Detection module}}.
            \item To finish and test the \textbf{\textit{Degradation Detection module}},
            organize the whole basic framework.
            \item Prepare and practice the Synthetic Data Method for the project,
            design the \textbf{\textit{Synthetic Data module}}. 
        \end{enumerate}
\end{itemize}

\noindent\rule{\textwidth}{0.4pt}
\centerline{\textbf{Complete basic framework at the finish of Degradation Detection Module}}
\noindent\rule{\textwidth}{0.4pt}

\begin{itemize}
    \item November
        \begin{enumerate}
            \item To finish and test the \textbf{\textit{Synthetic Data module}},
            figure out where to integrate it in the overall framework and
            how to ensure its compatibility with existing modules and test its performance.
            \item Prepare and practice the spatial analysis into the overall framework,
            study and practice deep learning methods, theoretically integrate them into the framework.
            \item Design, finish and test the \textbf{\textit{geographical weighted PCA module}},
            integrate into the established framework and interpret its results against the classic PCA results.
            \item Create new modules or extend existing modules to support spatial analysis and test them.
        \end{enumerate}
    \item December
        \begin{enumerate}
            \item Introduce deep learning methods into the existing framework,
            figure out how to integrate them into specific steps of modules and how to
            validate their performance and interpret their results.
            \item Integrate and adjust the whole modules, ensuring smooth interaction and data flow between them,
            like a software test suite.
            \item Finish 80\% of the first draft.
            \item Include the Satellite-derived data about wildfire into the project,
            explore the spatial analysis combined piecewise quantile regression using the
            established modules.
        \end{enumerate}
\end{itemize}

The below months are in 2026, each enumerated body in each month is the 
1st, 2nd, 3rd and 4th week of that month, '.' symbol means still to be determined.

\begin{itemize}
    \item January
        \begin{enumerate}
            \item Analyse the Satellite-derived data about wildfire and its relevance to
            the established methods, identifying potential improvements and adjustments.
            \item Finish the complete first draft.
            \item Make necessary revisions based on feedback and testing results.
            \item Conduct thorough testing and validation of the entire framework.
        \end{enumerate}
\end{itemize}

\noindent\rule{\textwidth}{0.4pt}
\centerline{\textbf{Complete First draft of the thesis}}
\noindent\rule{\textwidth}{0.4pt}

\begin{itemize}
    \item February
        \begin{enumerate}
            \item Finish the second draft of the thesis.
            \item .
            \item .
            \item Finish the complete thesis(third version).
        \end{enumerate}

    \item March
        \begin{enumerate}
            \item Preparing presentation materials and slides.
            \item Rehearsing the presentation and addressing potential questions.
            \item .
            \item .
        \end{enumerate}
\end{itemize}

\noindent\rule{\textwidth}{0.4pt}
\centerline{\textbf{Complete Second and final draft of the thesis}}
\noindent\rule{\textwidth}{0.4pt}


\begin{itemize}
    \item April
        \begin{enumerate}
            \item Presentation and defence of the thesis, determined by the coordinator.
            \item .
            \item .
            \item .
        \end{enumerate}
\end{itemize}

\noindent\rule{\textwidth}{0.4pt}
\centerline{\textbf{Presentation and defence}}
\noindent\rule{\textwidth}{0.4pt}




